Collaboratively Testing the Validity of Neuroscientific Models

### 2013-11-30 NeurounitFrontiers/Paper.tex
Revisions to sec 4

\section{A Collaborative Workflow}\label{sec:neuronunit_acitivities}
The \textit{NeuronUnit} package contains a collection of test families and associated capabilities relevant to neurophysiology, such as those described in the previous section. 
It is collaboratively developed on Github (\url{http://github.com/scidash/neuronunit}). 
Scientists can use this package to easily create test suites tailored to specific neuronal systems of interest. 
For example, scientists interested specifically in testing models of cells in the hippocampus would use \textit{NeuronUnit} to create tests in a GitHub repository containing an iPython notebook called \textit{hippocampus\_tests}, parameterized by data from the hippocampus.%, parameterizing them by relevant data from \textit{NeuroElectro}. 
%To make this easy, \textit{NeuronUnit} contains \emph{bridge functions} for retrieving data and generating tests from \textit{NeuroElectro} directly (Sec. \ref{sec:neuroelectro}).

@@

\section{Integrating With Neuroinformatics Tools and Infrastructure}\label{sec:neuronunit}
In the examples in the previous section, tests were generated from data explicitly, and models implemented capabilities directly. 
In many fields, however, both models and data are available from existing  repositories in standardized formats. 
In this section, we will show how to leverage this existing infrastructure to make parameterizing tests with published data and generating \textit{SciUnit} models from models implemented in some other manner (even in a language other than Python) nearly automatic. 
In particular, we will continue to focus on neurophysiology as our initial case study, using machine-readable models written in \textit{NeuroML} available from OpenSourceBrain (\textit{OSB}), and machine-readable data from \textit{NeuroElectro}.

\subsection{Reference Data for Tests from NeuroElectro}\label{sec:neuroelectro} 
\begin{figure}
\begin{python}
reference_data = neuroelectro.NeuroElectroSummary( # Pooled experimental data from  neuroelectro.org. 
  neuron = {'name':'Hippocampus CA1 pyramidal cell'}, # Neuron type according to NeuroLex ontology.  
  ephysprop = {'name':'Spike Width'}) # Electrophysiological property name in same.
reference_data.get_values()  # Get summary data for the above. 
model = models.OSBModel( # Initialize the model.
        'hippocampus', # Brain area.  
        'CA1_pyramidal_neuron', # Neuron type name in OSB.  
        'CA1PyramidalCell') # Model name (Migliore et al, 2005).
test = SpikeWidthTest( # Initialize the test.    
	reference_data = {'mean':reference_data.mean, 'std':reference_data.std}, # Summary statistics.
	model_args = {'current':40.0*pA}, # Somatic current injection in pA.  
	comparator = ZComparator), # A comparison class that implements a Z-score.  
record = test.judge(model) # (1) Check capabilities, (2) take the test, (3) generate a score and validate it, (4) bind the score to model/test combination. 
record.summarize() # Summarize the result.  
\end{python}
\vspace{-15px}
\caption{Working example of testing in \textit{NeuronUnit}. Imports excluded for clarity.}
\label{fig:neuronunit_example}

\end{figure}
The NeuroElectro project (\url{http://neuroelectro.org}) is an effort to make all published data on single cell neurophysiology available in a machine-readable format \citep{tripathy_neuroelectro:_2012}.  
Currently, up to 27 electrophysiological properties are available for 93 cell types, gathered from over 2000 single pieces of published data extracted from tables in published papers. 
We have made it easy to parameterize \textit{NeuronUnit} tests using data extracted from the NeuroElectro API. 
Tests can be based upon data from single journal articles, or from ensembles of articles with a common theme (e.g. about a particular neuron type). 
The former is illustrated in Figure \ref{fig:neuronunit_example}. 
Associated statistics of that data (e.g mean, standard error, and sample size) are attached and enable judgement of model output according to a chosen scoring mechanism. 
While NeuroElectro alone cannot judge all model aspects, it can serve to validate basic features of many neurophysiology models, such as resting membrane potential, action potential width, after-hyperpolarization amplitude, etc. 
As NeuroElectro is the only publicly curated source of such data, it represents a key component for \textit{NeuronUnit} test constuction.

\subsection{Models from NeuroML}\label{sec:neuroml_models}
NeuroML is a standardized model description language for neuroscience \citep{gleeson_neuroml:_2010}. 
It allows many neurophysiological and neuroanatomical models to be described in a simulator-independent fashion. 
Because NeuroML is an XML specification, model descriptions can be validated for correctness and queried for model properties and components. We use these facilities to expose capabilities automatically. 
It has proven ideal for model sharing, curation, and for answering both questions about \textit{what a model does} and \textit{how to do it} programatically.

NeuroConstruct (\cite{gleeson_neuroconstruct:_2007}, \url{http://www.neuroconstruct.org}) is a simulation manager that takes NeuroML models and hands off simulation to supported simulators. 
\textit{NeuronUnit} offers a \verbx{sciunit.Model} subclass called \verbx{NeuroConstructModel}, parameterized by a path to a NeuroML model.  
NeuroML can describe a wide range of models, but \verbx{NeuroConstructModel} currently supports only those that are time-integrable neurons with membrane potentials (exposing capabilities \verbx{TimeIntegrable} and \verbx{HasMembranePotential}). 
It can be instantiated directly or subclassed (useful if additional capabilities need to be manually exposed, for example) to generate a model suitable for testing (Fig. \ref{fig:ca1_model}).

@@

\begin{figure}
\begin{python}
class CA1PyramidalCellModel(NeuroConstructModel):
	"""CA1 Pyramidal Cell model from Open Source Brain."""
	def __init__(self,**kwargs):
		project_path = neuroconstruct.get_path("hippocampus","CA1_pyramidal_neuron","CA1PyramidalCell","neuroConstruct")
		NeuroConstructModel.__init__(self,project_path,**kwargs)
\end{python}
\vspace{-5px}
\caption{A model class corresponding to a CA1 Pyramidal Cell model from Open Source Brain}
\label{fig:ca1_model}
\vspace{-15px}
\end{figure}

>>>>

\section{A Collaborative Workflow}\label{sec:neuronunit_acitivities}
The \textit{NeuronUnit} package contains a collection of test families and associated capabilities relevant to neurophysiology, such as those described in the previous section. 
It is collaboratively developed on Github (\url{http://github.com/scidash/neuronunit}). 
Scientists can use this package to easily create test suites tailored to specific neuronal systems of interest. 
For example, scientists interested specifically in testing models of cells in the hippocampus would use \textit{NeuronUnit} to create tests in a GitHub repository containing an IPython notebook called \textit{hippocampus\_tests}, parameterized by data from the hippocampus.%, parameterizing them by relevant data from \textit{NeuroElectro}. 
%To make this easy, \textit{NeuronUnit} contains \emph{bridge functions} for retrieving data and generating tests from \textit{NeuroElectro} directly (Sec. \ref{sec:neuroelectro}).

@@

\section{Integrating With Neuroinformatics Tools and Infrastructure}\label{sec:neuronunit}
In the examples in the previous sections, tests were provided data explicitly, and models implemented capabilities directly. 
In many fields, however, both models and data are available from existing  community resources in standardized formats. 
In this section, we will show how to use this existing infrastructure to make parameterizing tests with published data and generating \textit{SciUnit} models from models implemented in some other manner (even in a language other than Python) nearly automatic. 
In particular, we will continue to focus on neurophysiology as our initial case study, using machine-readable models written in \textit{NeuroML} available from OpenSourceBrain (\textit{OSB}), and machine-readable data from \textit{NeuroElectro}.

\subsection{Reference Data for Tests from NeuroElectro}\label{sec:neuroelectro} 
\begin{figure}
\begin{python}
reference_data = neuroelectro.NeuroElectroSummary( # Pooled experimental data from  neuroelectro.org. 
  neuron = {'name':'Hippocampus CA1 pyramidal cell'}, # Neuron type according to NeuroLex ontology.  
  ephysprop = {'name':'Spike Width'}) # Electrophysiological property name in same.
reference_data.get_values()  # Get summary data for the above. 
model = models.OSBModel( # Initialize the model.
        'hippocampus', # Brain area.  
        'CA1_pyramidal_neuron', # Neuron type name in OSB.  
        'CA1PyramidalCell') # Model name (Migliore et al, 2005).
test = SpikeWidthTest( # Initialize the test.    
	reference_data = {'mean':reference_data.mean, 'std':reference_data.std}, # Summary statistics.
	model_args = {'current':40.0*pA}, # Somatic current injection in pA.  
	comparator = ZComparator), # A comparison class that implements a Z-score.  
score = test.judge(model)
\end{python}
\vspace{-15px}
\caption{Working example of testing in \textit{NeuronUnit}. Imports excluded for clarity.}
\label{fig:neuronunit_example}

\end{figure}
The NeuroElectro project (\url{http://neuroelectro.org}) is an effort to make all published data on single cell neurophysiology available in a machine-readable format \citep{tripathy_neuroelectro:_2012}.  
Currently, up to 27 electrophysiological properties are available for 93 cell types, gathered from over 2000 single pieces of published data extracted from tables in published papers. 
We have made it easy to parameterize the test families in \textit{Neuron\-Unit}  using data extracted from the NeuroElectro API. 
Tests can be based upon data from single journal articles, or from ensembles of articles with a common theme (e.g. about a particular neuron type). 
The latter is illustrated in Figure \ref{fig:neuronunit_example}, on lines 1-4. Once data has been retrieved from NeuroElectro, associated statistics (e.g. mean, standard error and sample size) can be extracted to parameterize tests (e.g. a test of spike widths on line 10). Data from NeuroElectro is can be used to validate a variety of basic electrophysiological features of neuron models, including spike width, resting membrane potential, after-hyperpolarization amplitude, etc. 
As NeuroElectro is the only publicly curated source of such data, it represents a key resource facilitating test construction.

\subsection{Models from NeuroML and OpenSourceBrain}\label{sec:neuroml_models}
NeuroML is a standardized model description language for neuroscience \citep{gleeson_neuroml:_2010}. 
It allows many neurophysiological and neuroanatomical models to be described in a simulator-independent fashion. Some simulators, notably the popular NEURON package\todo{cite NEURON}, can seamlessly export model specifications in the NeuroML format. 
Because NeuroML is an XML specification, model descriptions can be verified for correctness and queried for model properties and components. We leverage the latter facilities to expose model capabilities automatically. %NeuroML has proven ideal for model sharing, curation, and for answering both questions about \textit{what a model does} and \textit{how to do it} programatically.

NeuroConstruct (\cite{gleeson_neuroconstruct:_2007}, \url{http://www.neuroconstruct.org}) is a simulation manager that takes NeuroML models and hands off them off to a supported simulator for execution. 
\textit{NeuronUnit} provides a \verbx{sciunit.Model} subclass called \verbx{NeuroConstructModel} that takes a path to a NeuroML model as a parameter.  
NeuroML can describe a wide range of models, but our \verbx{NeuroConstructModel} class currently supports only  time-integrable point neurons (exposing capabilities \verbx{TimeIntegrable} and \verbx{HasMembranePotential} at minimum). 
It can be instantiated directly or subclassed (useful if additional capabilities need to be manually exposed, for example) to generate a model suitable for testing (Fig. \ref{fig:ca1_model}).

@@

\begin{figure}
\begin{python}
class CA1PyramidalCellModel(NeuroConstructModel):
	"""CA1 Pyramidal Cell model in the local file system."""
	def __init__(self):
		project_path = neuroconstruct.get_path("hippocampus","CA1_pyramidal_neuron","CA1PyramidalCell","neuroConstruct")
		NeuroConstructModel.__init__(self,project_path)
\end{python}
\vspace{-5px}
\caption{A model class corresponding to a CA1 Pyramidal Cell model from Open Source Brain}
\label{fig:ca1_model}
\vspace{-15px}
\end{figure}
